
---
title: 'Data603: Statistical Modelling with Data'
output:
  html_document:
    df_print: paged
---
<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
      line-height: 2em;
  }
td {  /* Table  */
  font-size: 12px;
}
.title {
  font-size: 38px;
  color: DarkRed;
}
 p {line-height: 2em;}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: Teal;
}
h3 { /* Header 3 */
  font-size: 22px;
  color: Green;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

&copy; Danika Lipman 2023, based on content created by Thuntida Ngamkham

# **MULTIPLE LINEAR REGRESSION PART 2 IN CLASS PROBLEMS**

### **In Class Practice Problem 5**

- For this problem we will be using the data from the **condominium.csv** file.
   a) Create a model with all interaction terms (based on the best model from last class)
   b) Conduct an appropriate test to determine if interaction terms should be included in the model.


```{r}
Condominium = read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/condominium.csv")
```

```{r}
reg <- lm(listprice~floors+livingarea+bedrooms+baths,data=Condominium)

smallReg<-lm(listprice~floors+livingarea+baths, data = Condominium)
summary(smallReg)
#0.965 for this model
```

```{r}

interactReg <- lm(listprice~(livingarea+floors+baths)^2, data=Condominium)
summary(interactReg)
```


----------------------------------------------------------------------

### **In Class Practice Problem 6**

- Data on last year's sale ($Y$ in 100,000s dollars) in 40 sales districts are given in the **sales.csv** file. This file also contains 
  + promotional expenditures ($X_1$: in 1,000s dollars), 
  + the number of active accounts ($X_2$), 
  + the number of competing brands ($X_3$), and
  + the district potential ($X_4$,coded) for each of the district
    + we will ignore this variable for now (do not include in the model) because it is not *quantitative*
  a) Find the best additive model using some or all of $X_1, X_2, X_3$
  b) Find the best interactive model using some or all of $X_1, X_2, X_3$
  c) Which of the two models would you use? 
  d) How will you interpret the coefficient for $X_3$? (Hint! there should be an interaction)

```{r}
Sales = read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/sales.csv")
```

```{r}
full = lm(Y~X1+X2+X3, data=Sales)

reduced = lm(Y~X2+X3, data=Sales)

Model1 = lm(Y~X2+X3+X2*X3, data=Sales)
summary(Model1)

Model2 = lm(Y~X1+X2+X3+X1*X3+X2*X3, data=Sales)
summary(Model2)
```

----------------------------------------------------------------------


### **In Class Practice Problem 7**

Suppose that we wish to investigate differences in credit card balance between marital status. Based on the Married variable, we can create a dummy variable which 0 is NO and 1 is YES. 

  a) Create a simple linear regression model to predict the credit card balance by using the Married variable. 
  b) What are the regression coefficients for this model? 
  c) How do you interpret the regression coefficients? **Ignore the individual t-test output**
  
```{r}
Credit=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/credit.csv")
```

```{r}
simpleLM <- lm(Balance~factor(Married), data=Credit)
summary(simpleLM)
```

----------------------------------------------------------------------

### **In Class Practice Problem 8**

- There is often curiosity and controversy surrounding professors’ salaries and whether they are overpaid or not paid enough. A university would like to study the effects of ranks and department on the salaries. 30 observations were randomly chosen from 3 different departments. The data are provided in **salary.csv** data file.
- Here is a description of the data:
  + Gender (0 = male, 1 = female)
  + rank ( 1= Asssistant, 2= Associate, 3 = Full)
  + dept: Deprtment (1 = Family Studies, 2= Biology, 3 = Business)
  + year: years since making rank
  + merit: average merit ranking
- The variable department has 3 levels which shall be dummy coded into 2 variables.
  + One called biology
  + One called business
- Rank also has 3 levels so we will code it into 2 dummy variables:
  + Assistant Professor
  + Full Professor
- Instead of the rank variable, practice how to interpret the dept variable.

```{r}
Salary = read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/salary.csv")
```

```{r}
sLM = lm(salary~factor(dept), data = Salary)
summary(sLM)
```

----------------------------------------------------------------------

### **In Class Practice Problem 9**

From the credit card example, use the lm() function to perform the best fit model. Recall the summary of the dataset:

- This Credit data set records:
  + Balance (average credit card debt for a number of individuals) 
  + Six quantitative predictors: age, cards (number of credit cards), education (years of education), income (in thousands of dollars), limit (credit limit), and rating (credit rating). 
  + Four qualitative variables: gender, student (student status), status (marital status), and ethnicity (Caucasian, African American or Asian). 
  
a) How would you interpret the regression coefficients (if possible)?
b) Would you recommend this model for predictive purpose?



```{r}
Credit=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/credit.csv")
```

```{r}
interlm = lm(Balance~Income+Limit+Age+Cards+Rating+Education+factor(Ethnicity)+factor(Gender)+factor(Student)+factor(Married),data=Credit)
summary(interlm)


test = lm(Balance~(Income+Student+Rating+Age+Limit+Cards), data=Credit)
summary(test)

best = lm(Balance~Income+Student+Rating+Age+Limit+Cards+Income*Rating+Income*Age+Income*Student+Limit*Rating+Limit*Student, data=Credit)
summary(best)

Realbest = lm(Balance~Income+Student+Rating+Age+Limit+Cards+Income*Rating+Income*Student+Limit*Rating+Limit*Student, data=Credit)
summary(Realbest)
```

```{r}
student = subset(Credit, Student == 'Yes')
studentLM = lm()

no = subset(Credit, Student == 'No')
Realbest = lm(Balance~Income+Student+Rating+Age+Limit+Cards+Income*Rating+Income*Student+Limit*Rating+Limit*Student, data=no)
summary(Realbest)
```
----------------------------------------------------------------------

### **Inclass Practice Problem 10**

Suppose you wanted to model the quality, $y$, of a product as a function of the pressure pounds per square inch (psi), at which it is produced. Four inspectors independently assign a quality score between 0 and 100 to each product, and then the quality, $y$, is calculated by averaging the four scores. An experiment is conducted by varying temperature in F. Fit a higher-order model to the data and sketch the scatterplot. The data are provided in **PRODQUAL.csv** file

```{r}
ProdQual = read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/PRODQUAL.csv")
```

```{r}
ggplot(data=ProdQual,mapping= aes(x=PRESSURE,y=QUALITY))+geom_point(color='green4')+ 
  geom_smooth()

quad = lm(QUALITY~PRESSURE+I(PRESSURE^2), data=ProdQual)
cube = lm(QUALITY~PRESSURE+I(PRESSURE^2)+I(PRESSURE^3), data=ProdQual)
quart = lm(QUALITY~PRESSURE+I(PRESSURE^2)+I(PRESSURE^3)+I(PRESSURE^4), data=ProdQual)

```


-------------------------------------------------------------------




# References
_-Gareth James & Daniela Witten & Trevor Hastie Robert Tibshirani, An Introduction to Statistical Learning with Applications in R: Springer New York Heidelberg Dordrecht London._

_-Wickham and Grolemund, R for Data Science: O'Reilly Media_


