---
title: "Untitled"
output: html_document
date: "2023-11-27"
---
---
title: 'Data603: Statistical Modelling with Data'
output:
  html_document:
    df_print: paged
---
<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
      line-height: 2em;
  }
td {  /* Table  */
  font-size: 12px;
}
.title {
  font-size: 38px;
  color: DarkRed;
}
 p {line-height: 2em;}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: Teal;
}
h3 { /* Header 3 */
  font-size: 22px;
  color: Green;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

&copy; Danika Lipman 2023, based on content created by Thuntida Ngamkham

# **EXPERIMENTAL DESIGN PART 2 In Class Problems**

```{r}
library(lmtest)

```

------------------------------------------------------------------

### **Inclass Practice Problem 33**

From the MVPC experiment, check Model Adequacy (Normality Assumption, Constant Variance Assumption, Independent error term Assumption )

```{r}
MVPC=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/MVPC.csv", header=TRUE)
library(lmtest)
```

```{r}
CRD=aov(Score~Treatment, data=MVPC) #Perform ANOVA for CRD
summary(CRD)

par(mfrow=c(2,2))
plot(CRD)

bartlett.test(Score~Treatment, data=MVPC)
shapiro.test(residuals(CRD))
bptest(CRD)
```

------------------------------------------------------------------

### **Inclass Practice Problem 34**

From the lifetime of AA battery experiment, check Model Adequacy (Normaillty Assumption, Constant Variance Assumption, Independent error term Assumption)

```{r}
LifetimeData = read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/lifetime.csv")
```

```{r}
CRD=aov(hrs~device, data=LifetimeData) #Perform ANOVA for CRD
summary(CRD)
```

------------------------------------------------------------------



