---
title: 'ASSIGNMENT 1: Multiple Linear Regression'
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---


## First order Model with Interaction Term (Quantitative Variables)

Deadline: Mar. 8, 2024, by 11:59 pm. Submit to Gradescope.ca

&copy; Thuntida Ngamkham (modified by Danika Lipman) 2024


**Problem 1**. (From Exercise 1) The amount of water used by the production facilities of a plant varies. Observations on water usage and other, possibility related, variables were collected for 250 months. The data are given in __water.csv file__. The explanatory variables are

TEMP= average monthly temperature (degree celsius)

PROD=amount of production (in hundreds of cubic)

DAYS=number of operationing day in the month (days)

HOUR=number of hours shut down for maintenance (hours)

The response variable is USAGE=monthly water usage (gallons/minute)

a. Fit the model containing all four independent variables. What is the estimated multiple regression equation?

b. Test the hypothesis for the full model i.e the test of overall significance. Use significance level 0.05. Ensure that you include the hypotheses, p-value, and conclusion.

c. Would you suggest the model in part b for predictive purposes? Which model or set of models would you suggest for predictive purposes? Hint: Use Individual Coefficients Test (t-test) to find the best model. Make sure you provide your hypotheses, p-value(s), and final model.)

d. Use Partial $F$ test to confirm that the independent variable (removed from part c) should be out of the model at significance level 0.05. Ensure that you include the hypotheses, p-value, and conclusion.

e. Obtain a 95% confidence interval of regression coefficient for TEMP from the model in part c. Give an interpretation. 

f. Use the method of Model Fit to calculate $R^2_{adj}$ and RSE to compare the full model in part a and the model in part c. Which model or set of models would you suggest for predictive purpose? For the final model, give an interpretation of $R^2_{adj}$ and RSE.

g. Build an interaction model to fit the multiple regression model from the model in part f. From the output, which model would you recommend for predictive purposes? Be sure to explain your process. 

```{r}
water <- read.csv("~/603/Ass1/water.csv")
Q1a = lm(USAGE~PROD+TEMP+HOUR+DAYS, data=water)
#SEAN WRITE IN THE FORMULA HERE Q1a

Q1c =  lm(USAGE~PROD+TEMP+HOUR, data=water)#removed days because it wasnt significant
Q1d = anova(Q1a, Q1c)

Q1e = confint(Q1c, parm = 'TEMP')

fullRsquared = summary(Q1a)$adj.r.squared
reducedRsquared = summary(Q1c)$adj.r.squared

Q1f = lm(USAGE~(PROD+TEMP+HOUR+DAYS)^2,data=water)
Q1fReduced = lm(USAGE~PROD+TEMP+HOUR+PROD*TEMP+PROD*HOUR, data=water)
```

**Answer To Question 1a** 

$$
\begin{aligned}
y &= 5.89 + 0.04X_1+ 0.17X_2 - 0.07X_3 - 0.02X_4 \epsilon\\
where\\
y &= \mbox{Usage}\\
X_1 &= \mbox{Prod}\\
X_2 &= \mbox{Temperature}\\
X_3 &= \mbox{Hour}\\
X_4 &= \mbox{Days}\\
\end{aligned}
$$


**Answer To Question 1b** The null hypothesis is that Temperature, production, days and hours all do not have an effect on the outcome, Usage.$$ H_0:PROD/DAYS/HOUR/TEMP = 0$$. This means the model does not explain any variance in the outcome, Usage. The alternate hypothesis is $$ H_1:PROD/DAYS/HOUR/TEMP \neq 0$$ However, after looking at the p values of the independent variables it can be concluded that all variables excluding Days have a significant relationship with Usage (p values are below 0.05). The p value of the model is 2x10^-16, which means at least one predictor significantly affects the outcome variable, Usage. 

  
**Answer to Question 1c** I would not suggest the model in part B for predictive purposes, because it includes an independent variable that does not significantly affect the outcome variable. The p value for Days is 0.502, which is not below our significance threshold of 0.05. I would use the model that omits Days, and uses Temperature, Hours, and production as predictors. The p values for these predictors are all below 0.05. 

**Prod** = p value is 2x10^-16
**Temp** = p value is 2x10^-16
**Hour** = p value is 4.23x10^-5
**Days** = p value is 0.502

The final model would look as this:

  $$ Usage = \beta_0+\beta_1*Prod+\beta_2*Temp+\beta_3*Hour$$

**Answer to Question 1d** The anova test shows a test stat of 0.4514, and a p value of 0.5023. This means that we do reject the null hypothesis in favour of our alternate.

H_0 = Reduced model and Full model have no significant difference 
H_1 = Reduced model has significantly better fit than the full model

Therefore we should drop the Days variable as an indicator.

**Answer to Question 1e** We are 95% confident the Temperature variable has an impact between 0.153, and 0.185. This means that Temperature usually has a positive relationship with the outcome variable.

**Answer to Question 1f** The $R^2_{adj} = 0.8867$ for our full model, and the RSE is 1.768. Our reduced model has an $R^2_{adj} = 0.8869$ and an RSE of 1.766. The higher adjusted R squared indicates the model's inputs more accurately describes the variance. The RSE being slightly lower indicates the reduced model may be less reliable than the full model. 

**Answer to Question 1g** I would recommend the interaction model that has interactions between PROD and HOUR, and PROD and TEMP. The $R^2_{adj} = 0.9651$ , which is noticeawbly higher than 0.8869 of the reduced additive model. 



------------------------------------------------------------


**Problem 2**. A collector of antique grandfather clocks sold at auction believes that the price received for the clocks depends on both the age of the clocks and the number of bidders at the auction. Thus, (s)he hypothesizes the first-order model


$$
\begin{aligned}
y &= \beta_0 + \beta_1X_1+\beta_2X_2 +\epsilon\\
where\\
y &= \mbox{Auction price (dollars)}\\
X_1 &= \mbox{Age of clock (years)}\\
X_2 &= \mbox{Number of bidders}
\end{aligned}
$$
A sample of 32 auction prices of grandfather clocks, along with their age and the number of bidders, is given in data file __GFCLOCKS.CSV__

a. Use the method of least squares to estimate the unknown parameters $\beta_0$, $\beta_1$, $\beta_2$ of the model.

b. Find the value of SSE that is minimized by the least squares method.

c. Estimate $s$, the standard deviation of the model, and interpret the result.

d. Find and interpret the adjusted coefficient of determination, $R^2_{Adj}$.

e. Construct the ANOVA table for the model and test the global F-test of the model at the  $\alpha$ = 0.05 level of significance. Be sure to state your hypotheses, p-values, and conclusion.

f. Test the hypothesis that the mean auction price of a clock changes as the number of bidders increases when age is held constant (i.e., when  $\beta_2\neq0$). (Use $\alpha$ = 0.05 )

g. Find a 95% confidence interval for $\beta_1$ and interpret the result.

h. Test the interaction term between the 2 variables at $\alpha = .05$. What model would you suggest to use for predicting y? Explain.

```{r}
GFCLOCKS <- read.csv("~/603/Ass1/GFCLOCKS.csv")

Q2a = lm(PRICE~AGE+NUMBIDS, data = GFCLOCKS)
Q2b = sum(Q2a$residuals^2)

#TO find s we need to solve for sqrt(MSE), where MSE = SSE/df
Q2c = sqrt(Q2b/29)
Q2d = summary(Q2a)$adj.r.squared

Q2e = anova(Q2a)
Q2f = summary(Q2a)
Q2g = confint(Q2a, "AGE")
Q2h = lm(PRICE~AGE+NUMBIDS+AGE*NUMBIDS, data = GFCLOCKS)
```

**Answer to Question 2a** $\beta_0 = -1338.95, \beta_1 = 12.74, \beta_2 = 85.95$. The equation for this model would be:
  $$ y = -1338.95+12.74X_1+85.95X_2+\epsilon$$

**Answer to Question 2b** `r Q2b`

**Answer to Question 2c** `r Q2c` This standard deviation means the actual values deivate from our prediction by about 133.4847 kilojoules per kilowatt-hour.

**Answer to Question 2d** Our $R^2_{Adj} = 0.8849$. This means our model is explaining the variance in prices well. This means there is a strong relationship between the predictors and the outcome variable, Price.

**Answer to Question 2e** `r Q2e` Our null hypothesis is: H0 = None of the predictors have an impact on the outcome variable. This means the model does not explain any of the variance. Our alternate hypothesis is H1 = One or more of the predictors have an impact on the outcome variable. This means the model is able to explain some of the variance seen. 

**Answer to Question 2f** The p value for NUMBIDS is 9.34x10^-11, which is below the significance threshold of 0.05. Therefore we can reject the null hypothesis and conclude that NUMBIDS has a significant positive impact on the price of the clock, while age is held constant. 

**Answer to Question 2g** The lower limit for our confidence interval for Age is 10.89, and the upper limit is 14.59. This means we are 95% confident the coefficient for Age is between these values. In clearer terms, we can say that age has a positive correlation with price,  increasing it between 10.89 and 14.59.

**Answer to Question 2h** I would use the interactive prediction model. This is because the adjusted R squared is higher compared to the additive model, (0.9489 versus 0.8849). The RSE is also lower for the interactive model (88.91 versus 133.5). The p value for the interaction term is below 0.05, meaning there is a significant interaction between Age and NUMBIDS affecting Price. 

---------------------------------------------------


**Problem 3**. __Cooling method for gas turbines.__  Refer to the Journal of Engineering for Gas Turbines and Power (January 2005) study of a high pressure inlet fogging method for a gas turbine engine. The heat rate (kilojoules per kilowatt per hour) was measured for each in a sample of 67 gas turbines augmented with high pressure inlet fogging. In addition, several other variables were measured, including cycle speed (revolutions per minute), inlet temperature (degree celsius), exhaust gas temperature (degree Celsius), cycle pressure ratio, and air mass flow rate (kilograms persecond). The data are saved in the __TURBINE.CSV__ file.

![The first and last five observations for the turbine data are listed in the table.]


(a) Write a first-order model for heat rate (y) as a function of speed, inlet temperature, exhaust temperature, cycle pressure ratio, and air flow rate.


(b) Test the overall significance of the model using $\alpha=0.01$ Be sure to state your hypotheses, p-values, and conclusion.


(c) Fit the best additive model to the data using the method of least squares. Test significance of predictors at $\alpha = 0.06$.Be sure to state your hypotheses, p-values, and conclusion.

(d)  Test all possible interaction terms for the best model in part (c) at $\alpha = .06$. What is the final model would you suggest to use for predicting y? Explain.

(e) Give practical interpretations of the $\beta_i$ estimates. 

(f) Find RSE, $s$ from the model in part (d)

(g) Find the adjusted-R2 value from the model in part (d) and interpret it.

(h) Predict a heat rate (y) when a cycle of speed = 273,145 revolutions per minute,  inlet temperature= 1240 degree celsius, exhaust temperature=920 degree celsius, cycle pressure ratio=10 kilograms persecond, and air flow rate=25 kilograms persecond.

```{r}
TURBINE <- read.csv("~/603/Ass1/TURBINE.csv")
Q3a = lm(HEATRATE~RPM+INLET.TEMP+EXH.TEMP+CPRATIO+AIRFLOW, data=TURBINE)
#B is CPRATIO and AIRFLOW not sig
Q3c = lm(HEATRATE~RPM+INLET.TEMP+EXH.TEMP, data=TURBINE)

Q3d = lm(HEATRATE~(RPM+INLET.TEMP+EXH.TEMP)^2, data=TURBINE)
Q3d2 = lm(HEATRATE~(RPM+INLET.TEMP+EXH.TEMP+RPM*INLET.TEMP+RPM*EXH.TEMP), data=TURBINE)

```

**Answer to Q3a**
$$
\begin{aligned}
y &= 13,610 + 0.088X_1- 9.2X_2 + 14.39X_3 + 0.35X_4 - 0.85X_5 \epsilon\\
where\\
y &= \mbox{Heat Rate (dollars)}\\
X_1 &= \mbox{RPM (years)}\\
X_2 &= \mbox{Inlet Temperature}\\
X_3 &= \mbox{Exhaust Temperature}\\
X_4 &= \mbox{Cycle Pressure Ratio}\\
X_5 &= \mbox{Airflow}\\
\end{aligned}
$$


**Answer to Q3b** The null hypothesis is $$ H_0:RPM/InletTemp/ExhuastTemp/CPRatio/Airflow = 0$$. This means the predictors have no influence on the outcome variables. The alternate hypothesis is $$ H_0:RPM/InletTemp/ExhuastTemp/CPRatio/Airflow \neq 0$$. This means one or more predictor has an influence on the outcome, Heatrate.  The pvalues of all variables except for CPRatio and Airflow are below the threshold of 0.01. 

**RPM** =  p value of 2.64x10^-8

**InletTemp**=  p value of 6.86x10^-8

**ExhaustTemp** =  p value of 0.000102

**CPRatio** =  p value of 0.9905

**Airflow**=  p value of 0.0598

These p values mean that we are able to say Airflow and CPRatio do not significantly contribute to the model, and should not be used as predictors. 


This means that CRPratio and Airflow do not contribute to the model and do not serve predictive purposes. We are also able to reject our null hypothesis in favor of our alternate, since one or more predictors had a significant impact on the outcome, Heatrate.

**Answer to Q3c** Our null hypothesis is $$ H_0:RPM/InletTemp/ExhuastTemp= 0$$ The equation of the best additive model is:
$$
\begin{aligned}
y &= 14,360 + RPMX_1+ InletTempX_2 - ExhaustTempX_3  \epsilon\\
\end{aligned}

$$
The p values for the best additive model were:

**RPM** = 2.55x10^-14

**InletTemperature** = 2x10^-16

**ExhaustTemperature** = 1.06x10^-7

This means that all of these terms have a significant impact on the Heatrate. We can therefore reject our null hypothesis in favor of our alternate hypothesis. 

**Answer to Q3d** The final interactive model equation is: 

$$ Heatrate = \beta_0+\beta_1*RPM+\beta_2*InletTemp+\beta_3*Exh$$

**Answer to Q3e**

**Answer to Q3f**

**Answer to Q3g**

**Answer to Q3h**


