
---
title: 'Data603: Statistical Modelling with Data'
output:
  html_document:
    df_print: paged
---
<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
      line-height: 2em;
  }
td {  /* Table  */
  font-size: 12px;
}
.title {
  font-size: 38px;
  color: DarkRed;
}
 p {line-height: 2em;}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: Teal;
}
h3 { /* Header 3 */
  font-size: 22px;
  color: Green;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

&copy; Danika Lipman 2023, based on content created by Thuntida Ngamkham

# **MULTIPLE LINEAR REGRESSION PART 4 IN CLASS PROBLEMS**

```{r}
library(ggplot2)
library(lmtest)
library(mctest)
library(car)
library(MASS)
```

------------------------------------------------------------------

### **Inclass Practice Problem 17**

From the clerical staff work hours, use residual plots to conduct a residual analysis of the data.  Check Linearity Assumption.  If a trend is detected, how would you like to transform the predictors in the model?

- Begin with the model Y ~ X2 + X4 + X5

```{r}
Clerical=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/CLERICAL.csv")
```

```{r}
fullmodel = lm(Y~X2+X4+X5, data=Clerical)

ggplot(fullmodel, aes(x=.fitted, y=.resid)) +
  geom_point() +geom_smooth()+
  geom_hline(yintercept = 0) 

ggpairs(Clerical,lower = list(continuous = "smooth_loess", combo =
  "facethist", discrete = "facetbar", na = "na"))

highmodel = lm(Y~X2+X4+X5+I(X5^2), data=Clerical)
summary(highmodel)

ggplot(highmodel, aes(x=.fitted, y=.resid)) +
  geom_point() +geom_smooth()+
  geom_hline(yintercept = 0) 
```

------------------------------------------------------------------

### **In Class Problem 18**

From the clerical staff work hours, use residual plots to conduct a residual analysis of the data.  Check Equal Variance Assumption by graphs and the Breusch-Pagan test.  If you detect a trend, how would you like to transform the predictors in the model?

- Begin with the model Y ~ X2 +I(X2^2) + X4 + X5

```{r}
Clerical=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/CLERICAL.csv")
```

```{r}
highmodel = lm(Y~X2+X4+X5+I(X2^2), data=Clerical)
summary(highmodel)

bptest(highmodel) #p value of 0.152 so it is common variance

ggplot(highmodel, aes(x=.fitted, y=.resid)) +
  geom_point(colour = "purple") +
  geom_hline(yintercept = 0) +
  geom_smooth(colour = "green4")+
  ggtitle("Residual plot: Residual vs Fitted values")  

#a scale location plot
ggplot(highmodel, aes(x=.fitted, y=sqrt(abs(.stdresid)))) +
  geom_point(colour = "purple") +
  geom_hline(yintercept = 0) +
  geom_smooth( colour = "green4")+
   ggtitle("Scale-Location plot : Standardized Residual vs Fitted values") 
```


------------------------------------------------------------------


### **In Class Problem 19**

From the clerical staff work hours, use residual plots to conduct a residual analysis of the data.  
- Check Normality Assumption by graphs and the Shapiro-Wilk normality test. 
- If you detect a trend, how would you like to transform the predictors in the model?
- Begin with the model Y ~ X2 +I(X2^2) + X4 + X5

```{r}
Clerical=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/CLERICAL.csv")
```

```{r}
highmodel = lm(Y~X2+X4+X5+I(X2^2), data=Clerical)
summary(highmodel)

ggplot(Clerical, aes(sample=highmodel$residuals)) +
  stat_qq() +
  stat_qq_line()

shapiro.test(residuals(highmodel))
```


------------------------------------------------------------------


### **In Class Problem 20**

- From the credit card example, check for Multicollinearity  by scatter plots between independent predictors and VIF test .  
  + Note! consider only main effect predictors
  + Use the model BALANCE ~ Income + Rating + Age + Limit + Cards + factor(Student)
  
  
```{r}
Credit=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/credit.csv")

```

```{r}
model = lm(Balance~Income + Rating + Age + Limit + Cards + factor(Student), data = Credit)
vif(model)
```


------------------------------------------------------------------


### **In Class Problem 21**

- From the credit card example, after dropping the Limit variable, find the best additive model and check for Multicollinearity by scatter plots between independent predictors and VIF test. 
  + Note! consider only main effect predictors
  
  
```{r}
Credit=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/credit.csv")
```

```{r}
model = lm(Balance ~ Income + Rating + Age + Cards + factor(Student), data=Credit)
summary(model)

model2 = lm(Balance ~ Income + Rating + Age + factor(Student), data=Credit)
summary(model2)

vif(model2)
```
------------------------------------------------------------------


### **In Class Problem 22**

- From the clerical staff work hours, using residual plots to conduct a residual analysis of the data.  Check any potential outliers.   
- use the model Y ~ X2 + I(X2^2) + X4 + X5
  
  
```{r}
Clerical=read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/CLERICAL.csv")
```

```{r}
model=lm(Y ~ X2 + I(X2^2) + X4 + X5, data =Clerical)
summary(model)

#Residual vs fit
ggplot(model, aes(x=.fitted, y=.resid)) +
  geom_point() +geom_smooth()+
  geom_hline(yintercept = 0) 

#Rsidual vs leverage
plot(model, which = 5)

#normal QQ plot
ggplot(Clreical), aes(sample=model$residuals)) +
  stat_qq() +
  stat_qq_line()

#outliers
lev=hatvalues(model)
p = length(coef(model))
n = nrow(Clerical)
outlier2p = lev[lev>(2*p/n)]
outlier3p = lev[lev>(3*p/n)]
print(outlier3p)


#Cooks distance
Clerical[cooks.distance(model)>0.5,] 

#Colinerality says colineraity in 
vifmodel = lm(Y~X4+X5, data=Clerical)
vif(vifmodel)

#Scale location
#a scale location plot
ggplot(model, aes(x=.fitted, y=sqrt(abs(.stdresid)))) +
  geom_point(colour = "purple") +
  geom_hline(yintercept = 0) +
  geom_smooth( colour = "green4")+
   ggtitle("Scale-Location plot : Standardized Residual vs Fitted values") 
```

------------------------------------------------------------------

### **In Class Problem 23**

- Check assumptions for the model below to to predict executive salary (Y)

$$
Y=\beta_0+\beta_1X_1+\beta_2X_1^2+\beta_3X_2+\beta_4X_3+\beta_5X_4+\beta_6X_5+\beta_7X_3*X_4+\epsilon 
$$
  
  
```{r}
ExecSal = read.csv("https://raw.githubusercontent.com/DanikaLipman/DATA603/main/EXECSAL2.csv")
```


------------------------------------------------------------------



# References
_-Gareth James & Daniela Witten & Trevor Hastie Robert Tibshirani, An Introduction to Statistical Learning with Applications in R: Springer New York Heidelberg Dordrecht London._

_-Wickham and Grolemund, R for Data Science: O'Reilly Media_


